
## Loading and preprocessing the data
```{r load_data, results='hide'}
data_fname <- "activity.csv"
zipfile <- "activity.zip"
if (!file.exists(data_fname)) {
    unzip(zipfile)
    }
data <- read.csv(data_fname)
options(scipen = 1, digits = 2)
```

## What is mean total number of steps taken per day?
```{r mean_steps}
is_na <- is.na(data$steps)
data1 <- data[!is_na,]
data2 <- aggregate(data1$steps, by = list(data1$date), FUN = "sum")
names(data2) <- c("date", "steps")
data2$date <- as.Date(data2$date)
main <- "Histogram of steps per day"
xlab <- "Total number of steps"
hist(data2$steps, col = "red", main = main, xlab = xlab, breaks = 10)
mean <- mean(data2$steps)
median <- median(data2$steps)
```
Mean number of steps per day is `r mean`.

Median number of seps per day is `r median`.

## What is the average daily activity pattern?
```{r daily_activity}
data3 <- aggregate(data1$steps, by = list(data1$interval), FUN = "ave")
names(data3) <- c("interval", "steps_ave")
data3$steps_ave <- data3$steps_ave[,1]
xlab = "5-minute interval"
ylab = "Average number of steps"
plot(data3$interval, data3$steps_ave, type = "l", xlab = xlab, ylab = ylab)
pmax <- which.max(data3[,2])
interval <- data3$interval[pmax]
```
5-minute interval '`r interval`' contains the maximum number of steps on average.

## Imputing missing values
```{r impute_missing_values}
nas <- dim(data)[1] - dim(data1)[1]
data4 <- data.frame(date = data$date, interval = data$interval, steps <- data$steps)
for (i in seq(is_na)) {
    if (is_na[i]) {
        interval <- data4$interval[i]
        ind <- match(interval, data3$interval)
        data4$steps[i] <- as.integer(data3$steps[ind])
    }
}
data5 <- aggregate(data4$steps, by = list(data4$date), FUN = "sum")
names(data5) <- c("date", "steps")
main <- "Histogram of steps per day (filled dataset)"
xlab <- "Total number of steps"
hist(data5$steps, col = "red", main = main, xlab = xlab, breaks = 10)
mean2 <- mean(data5$steps)
median2 <- median(data5$steps)
mean_change <- 100 * abs(mean - mean2) / mean
median_change <- 100 * abs(median - median2) / median
diff1 <- 100 * abs(mean - median) / mean
diff2 <- 100 * abs(mean2 - median2) / mean2
```
Total number of missing values: `r nas`.

To replace missing values (NA), I choose average for that 5-minute interval across all days.

Mean number of steps per day in filled dataset is `r mean2`.

Median number of steps per day in filled dataset is `r median2`.

After filling mean value has changed by `r mean_change` (%).

Median value has changed by `r median_change` (%).

Before filling median-mean difference was `r diff1` (%).

After filling median-mean difference became `r diff2` (%).

## Are there differences in activity patterns between weekdays and weekends?
```{r pattern_differences, fig.height=10}
data4$date <- as.Date(data4$date)
days <- vector("character", length = length(data4$date))
for (i in seq(days)) {
    day <- weekdays(data4$date[i])
    if (day == "суббота" || day == "воскресенье") {
        days[i] <- "weekend"
    }
    else {
        days[i] <- "weekday"
    }
}
data4$days <- days
data_weekday <- data4[data4$days == "weekday",]
data_weekend <- data4[data4$days == "weekend",]
data6 <- aggregate(data_weekday$steps, by = list(data_weekday$interval), FUN = "mean")
names(data6) <- c("interval", "steps_ave")
xlab = "5-minute interval"
ylab = "Average number of steps"
data7 <- aggregate(data_weekend$steps, by = list(data_weekend$interval), FUN = "mean")
names(data7) <- c("interval", "steps_ave")
par(mfrow = c(2, 1))
ttl = "weekday"
plot(data6$interval, data6$steps_ave, type = "l", xlab = xlab, ylab = ylab, main = ttl)
ttl = "weekend"
plot(data7$interval, data7$steps_ave, type = "l", xlab = xlab, ylab = ylab, main = ttl)
```